SPP-Perc: Peceptron Learning Based Enhanced Signature Path Prefetcher

### 2018-12-04 design_arch.tex
Modified sections 4.2 and the inference section in 3.1

\pg{this discussion below has redundancy and doesn't flow well.
  Rework it}

The perceptron sum can be considered the sum of the individual
contribution of a feature.  The value of each contribution corresponds
to a feature's confidence in the final decision on whether to prefetch
a cache line. By summing the individual contributions, the final
perceptron sum denotes the overall confidence for that prefetch
suggestion.  By comparing the perceptron sum to two different
thresholds, we divide the confidence scale of the prefetch suggestion
into three bins. The first bin corresponds to the lowest confidence
and leads to prefetch candidate being rejected.  The next bin
corresponds to prefetches with a moderate confidence level.  This bin
prefetches directeky to fill the bigger last level cache and prevents
polluting the smaller L2 Cache.  The highest confidence prefetches are
filled in the L2 cache.

The sum of the perceptron weights are also used to decide whether to
place the data in the L2 or the last level cache.  All prefetch
candidates that are qualified until this stage are compared against
PERC\_THRESHOLD\_HI to decide the fill level. The two thresholds:
PERC\_THRESHOLD\_LO and PERC \_THRESHOLD\_HI are empirically set.

>>>>

%\pg{this discussion below has redundancy and doesn't flow well.
%  Rework it}

Each feature's confidence contributes to the final decision on 
whether to prefetch a cache line, with the summation of all features' 
confidences dentoting the overall confidence for a prefetch. 
The final sum is compared against two thresholds: 
 PERC\_THRESHOLD\_HI and  PERC\_ THRESHOLD\_ LO to decide the
 placement of the data in the cache heirarchy. Prefetches that 
have a confidence above the higher threshold are placed into 
the L2 cache since the observed features strongly recommend  
the prefetch. A prefetch that results in a confidence between 
PERC\_ THRESHOLD\_ LO and PERC\_ THRESHOLD\_ HI are placed in 
the larger last level cache, as the filter is moderately confident of 
the cache line's use, but seeks to reduce pollution in the smaller 
L2 cache. Finally, the remaining prefetches that are below the 
lower threshold are not prefetched due to a lack of correlation 
between the current observed features and the predicction from 
the base prefetch engine.

%GAC: The section's previous contents are below, in case the above 
%rewritten work is not what we were looking for in terms of a rewording
%
%The perceptron sum can be considered the sum of the individual
%contribution of a feature.  The value of each contribution corresponds
%to a feature's confidence in the final decision on whether to prefetch
%a cache line. By summing the individual contributions, the final
%perceptron sum denotes the overall confidence for that prefetch
%suggestion.  By comparing the perceptron sum to two different
%thresholds, we divide the confidence scale of the prefetch suggestion
%into three bins. The first bin corresponds to the lowest confidence
%and leads to prefetch candidate being rejected.  The next bin
%corresponds to prefetches with a moderate confidence level.  This bin
%prefetches directeky to fill the bigger last level cache and prevents
%polluting the smaller L2 Cache.  The highest confidence prefetches are
%filled in the L2 cache.
%
%
%The sum of the perceptron weights are also used to decide whether to
%place the data in the L2 or the last level cache.  All prefetch
%candidates that are qualified until this stage are compared against
%PERC\_THRESHOLD\_HI to decide the fill level. The two thresholds:
%PERC\_THRESHOLD\_\_LO and PERC \_THRESHOLD\_HI are empirically set.

### 2018-12-04 design_impl.tex
Modified sections 4.2 and the inference section in 3.1

\begin{itemize}
\item \textbf{Base Address} of the demand access that triggered the
  prefetch.  Since prefetches are triggered from the L2 demand access,
  they tend to correlate to the triggering address.

@@

\item \textbf{Program Counter XOR Depth}: The PC is for the
  instruction that triggered the prefetch chain.  Depth refers to the
  iteration count of the lookahead stages.  In general, the PC is not
  considered as a good basis for doing lookahead prefetching as all
  the prefetches with depth >= 1 are aliased into the same PC which
  would not be true in an actual demand access.  This feature resolves
  a PC into a different value for each lookahead depth of prefetch
  speculation, giving a more accurate correlation in lookahead cases.
  This is akin to the concept of Virtual Program Counters~\cite{VPC}
  introduced by Kim \textit{et. al.} for indirect branch prediction.

@@

\item \textbf{Program Counter XOR Delta}: This feature tells us if a
  given PC favours particular value(s) of delta.  As noted earlier,
  while the PC in itself does not convey much useful information, this
  hash resolves the PC into different values based on the tendency of
  that PC to favour a certain delta.  Thus, the dynamic nature of
  different instances of the same memory access instruction can be
  captured here.

\item \textbf{Confidence}: The integer confidence on a scale of 0 to
  100 that was used in the original SPP design. While the original
  confidence is not used directly for decision making, it can still
  contribute to the final decision made.

@@

% djimenez: this is a little vague. can you be more precise?
\item \textbf{Current Signature XOR Delta}: Recall from the discussion
  of SPP in Section \ref{Background-SPP} that the new signature is
  generated using the old signature and the current block delta.
%Gino: Added the sentence below, is this correct?
  The result of this feature is the next delta that is predicted to be
  accessed based on the delta predicted by SPP. While ``Current
  Signature XOR Delta'' is not the exact formula for generating the
  future signature, it gives an approximate idea of the path that the
  combination of these two values can lead to.

>>>>

\begin{itemize}
\item \textbf{Base Address}: The base address of the demand 
access that triggers the prefetch coresponds to a stream of 
future accesses that SPP and the PPF have seen before. 
Therefore, PPF will correlate the past behavior of this address 
to the confidence of a prefetch.

@@

\item \textbf{Program Counter XOR Depth}: The PC is for the
  instruction that triggered the prefetch chain.  Depth refers to the
  iteration count of the lookahead stages.  In general, the PC is not
  considered as a good basis for doing lookahead prefetching as all
  the prefetches with depth >= 1 are aliased into the same PC which
  is not true in an actual demand access.  This feature resolves
  a PC into a different value for each lookahead depth of prefetch
  speculation, giving a more accurate correlation in lookahead cases.
  This is akin to the concept of Virtual Program Counters~\cite{VPC}
  introduced by Kim \textit{et. al.} for indirect branch prediction.

@@

\item \textbf{Program Counter XOR Delta}: This feature tells us if a
  given PC favours particular value(s) of delta.  As noted earlier,
  while the PC alone does not convey useful information, this
  hash resolves the PC into different values based on the tendency of
  that PC to favour a certain delta.  Thus, the dynamic nature of
  different instances of the same memory access instruction is 
  captured here.
 % GAC: Seems more predictable than dynamic if its capturing 
 % what is essentially a streaming access pattern. Not quite sure how to rephrase,
 % if I'm correct

 % GAC: Again having trouble phrasing the below point. I was attempting to say
 % that despite SPP having low confidence due to speculation, PPF may 
 % reinforce prefetches with lower confidence if SPP has had a tendancy to be 
 % correct in the past.
\item \textbf{Confidence}: The integer confidence, on a scale of 0 to
  100, that was used in the original SPP design. While the original 
  confidence does not directly make the decision to prefetch,
  PPF correlates it to the correctness of a proposed prefetch.
  While the original SPP may have dismissed a prefetch due to running
  further into speculation, PPF can use the original confidence as indicator
  of not only when prefetches become less confident, but also how likely
  a low confident speculation is correct in the context of other features.

@@

\item \textbf{Current Signature XOR Delta}: Recall from the discussion
  of SPP in Section \ref{Background-SPP} that the new signature is
  generated using the old signature and the current block delta.
%Gino: Added the sentence below, is this correct?
  The result of this feature is the next delta that is predicted to be
  accessed based on the delta predicted by SPP. While ``Current
  Signature XOR Delta'' is not the exact formula for generating the
  future signature, it gives an approximate idea of the path that the
  combination of these two values can lead to.
